### 2020-04-08 draft.tex
latest

\section{Introduction}
A polynomial functor $P : \Set \to \Set$ is a sum of representables
\begin{equation}\label{eqn.poly}
P(X) := \sum_{b \in B} X^{E_b}
\end{equation}
where the family of sets $E_b$ depends on $b\in B$. This data is known
in the computer science literature as a ``container'' \cite{Abbott}, but since such an indexed
family of sets can be represented by as a function
\[
  \begin{tikzcd}
    E \arrow[d, "\pi"] \\
    B
  \end{tikzcd}
\]
we will refer to it as a \emph{bundle}.

@@

\begin{thm}[cite]
The category of polynomial functors and natural transformations is equivalent to
the category of bundles and contravariant morphisms.
\end{thm}

@@

We will show in this paper that Dirichlet functors are, quite robustly, the
contravariant analogue of polynomial functors. In particular, the many equivalent ways to
say that a functor is polynomial have contravariant analogues.
\begin{thm}[cite]\label{thm:polynomial.set.characterization}
  Let $P : \Set \to \Set$ be a functor. Then the following are equivalent.
  \begin{enumerate}
  \item $P$ is polynomial.
  \item $P$ is the sum of covariant representables.
  \item There is a bundle $\pi : E \to B$ together with a natural isomorphism
    $$P(X) \cong \sum_{b \in B} X^{E_b}.$$
    Or, equivalently, a natural isomorphism of $P$ with the composite
    $$\Set \xto{{\Delta_{!_E}}} \Set_{/E} \xto{{\Pi_\pi}} \Set_{/B} \xto{{\Sigma_{!_B}}} \Set.$$
  \item $P$ is accessible and preserves connected limits.
  \end{enumerate}
\end{thm}

@@

Now, object-wise, a Dirichlet functor and a polynomial functor are determined by
the same data --- a bundle $\pi : E \to B$ of sets. Accordingly, one would
expect for any set $N$ a transformation
$$X^N \mapsto N^X$$
turning polynomial functors into Dirichlet functors, and vice versa. But the
natural transformations between each sort of functor induce different morphisms
between the bundles; natural transformations between polynomial functors induce
contravariant bundle morphisms, while natural transformations between Dirichlet
functors induce covariant bundle morphisms. However, if we restrict to those
morphisms of bundles which are \emph{isovariant} on the fibers --- that is, the
pullback diagrams of the form
\[
  \begin{tikzcd}[column sep=large]
E \arrow[r, "\term{tot}(f_{\sharp})"] \arrow[d, "\pi"'] \arrow[dr, phantom,
"\ulcorner" very near start] & E' \arrow[d, "\pi'"] \\
B \arrow[r, "f"']                                       & B'                  
\end{tikzcd}
\]
which preserve the number of elements in each row --- we find that such a morphism
is both a co- and contravariant morphism of bundles. It is well known that such
\emph{cartesian} morphisms of bundles correspond to \emph{cartesian} natural
transformations between polynomial functors \cite[Theorem 3.8]{GambinoKock} --- those whose naturality
squares are pullbacks. This is true as well for Dirichlet functors.
\begin{thm}
A natural transformation $D \to D'$ of Dirichlet functors is Cartesian if and
only if the corresponding bundle map
\[
  \begin{tikzcd}[column sep=large]
E \arrow[r, "\term{tot}(f_{\sharp})"] \arrow[d, "\pi"'] \arrow[dr, phantom,
"\ulcorner" very near start] & E' \arrow[d, "\pi'"] \\
B \arrow[r, "f"']                                       & B'                  
\end{tikzcd}
\]
is a pullback. As a corollary, we have an equivalence of categories
$$\type{Poly}_{\ulcorner} \simeq \Dir_{\ulcorner}$$
between polynomial functors with cartesian natural transfromations and Dirichlet
functors with cartesian natural transfromations.
\end{thm}

@@

\begin{defn}
  We let $\downarrow$ be the walking arrow --- the category $\term{dom} \to
  \term{cod}$ consisting of a single morphism. We denote by $\sigma : \downarrow
  \to (\downarrow)\op$ the self-dualizing isomorphism of $\downarrow$, and note
  that $\sigma\inv = \sigma\op$.
\end{defn}

@@

The existence of the final adjoint can be deduced from the fact that $!^{(-)}$ preserves all
limits. It sends a bundle $\pi : E \to B$ to the largest subset $\term{ZC}(B)$ of $B$ for
which the following square is a pullback
\[
  \begin{tikzcd}[column sep=large]
0 \arrow[r] \arrow[d] \arrow[dr, phantom,
"\ulcorner" very near start] & E \arrow[d, "\pi"] \\
\term{ZC}(\pi) \arrow[r, hookrightarrow]                                       & B
\end{tikzcd}
\qedhere
\]
\end{proof}
We prove a few quick facts we will use later.
\begin{lem}\label{lem:bang.preserves.connected.colimits}
  The functor $!_{(-)}\colon\Set\to\Set^{\down}$ from \eqref{eqn.bangbang} preserves connected colimits.
\end{lem}
\begin{proof}
  To see that $!_{(-)}$ preserves connected colimits, recall that colimits in $\Set^{\down}$ are calculated pointwise. It remains to show, then, that the map of bundles
  \[
    \begin{tikzcd}
    \colim X_i \arrow[r, equals] \arrow[d] & \colim X_i \arrow[d] \\
    \colim 1 \arrow[r] & 1 
    \end{tikzcd}
  \]
  is an isomorphism in $\Set^{\down}$, for which it suffices to show that
  $\colim 1$ is terminal. But the colimit of a diagram of terminal objects is the set
  of connected components of its indexing category. Since we assumed the
  indexing category was connected, this contains a single element.
\end{proof}

@@

\begin{rmk}
The six adjoints of \cref{prop:adjoint.sextuple} correspond, under
the equivalence of \cref{thm:dirichlet.set.characterization}, to
\begin{align*}
  \term{ZC}(D) &\sim \mbox{the coefficient of $0^X$ in $D$.}\\
  !^{C} &\sim X \mapsto C \times 0^X \\
  \cod(D) &\sim D(0) \\
  \const(C) &\sim X \mapsto C \\
  \dom(D) &\sim D(1) \\
  !_{C} &\sim X \mapsto C^X 
\end{align*}
\dis{If we're using \cref{thm:dirichlet.set.characterization}, shouldn't we be thinking in terms of bundles?} In particular, $!_{(-)}$ corresponds to the Yoneda embedding.
\end{rmk}

@@

\begin{thm}\label{thm:set.poly.cart.equiv.dir.cart}
  There is an equivalence of categories
  $$\type{Poly}_{\ulcorner} \simeq \Dir_{\ulcorner}$$
  between the category of polynomial functors and cartesian transformations and
  Dirichlet functors and cartesian transformations. This equivlalence sends
  representables to representables
  $$(-)^N \mapsto N^{(-)}.$$
\end{thm}
\begin{proof}
This follows by composing the equivalence of $\Dir_{\ulcorner}$ with
$\Set^{\down}$ from
\cref{prop:set.dirichlet.cartesian.iff.bundle.cartesian} with that of
\cite[Proposition 3.14]{GambinoKock}, noting that
$\Set^{\down}_{\ulcorner}$ is the category of $(1,1)$-polynomials. 
\end{proof}

\begin{cor}
  Let $D$ be a Dirichlet functor. Then the category
  $\Dir_{\ulcorner/D}$
  of Dirichlet functors with a cartesian map to $D$ is a topos.
\end{cor}
\begin{proof}
By Theorem \ref{thm:set.poly.cart.equiv.dir.cart}, this category is equivalent
to $\type{Poly}_{\ulcorner/P}$ for a polynomial $P$. But this is a topos as
observed in Remark 2.6.2 of \cite{GHK:Analytic.Monads}.
\end{proof}

@@

  with left exact left adjoint given by the vertical / cartesian factorization system:
  \[
\begin{tikzcd}
E' \arrow[r] \arrow[d] & E \arrow[d, "\pi"] \\
B' \arrow[r]           & B                 
\end{tikzcd} \quad\quad\mapsto\quad\quad
\begin{tikzcd}
E' \arrow[d] \arrow[r, dashed]    & \bullet \arrow[d] \arrow[r] \arrow[rd, phantom, "\ulcorner" very near start] & E \arrow[d, "\pi"] \\
B' \arrow[r, equals] & B' \arrow[r]                           & B                 
\end{tikzcd}
  \]
  As a corollary, $\Dir_{\ulcorner/D} \hookrightarrow \Dir_{D}$ is a
  subtopos inclusion.
\end{thm}
\begin{proof}
We have displayed both the action of the left localization and its unit. The
counit is always an isomorphism since pullbacks are unique up to unique
isomorphism.

>>>>

\section{Introduction}
A polynomial functor $P : \Set \to \Set$ is a sum of representables
\begin{equation}\label{eqn.poly}
P(X) := \sum_{b \in B} X^{E_b}
\end{equation}
where the family of sets $E_b$ depends on $b\in B$. This data is known
in the computer science literature as a ``container'' \cites{abbott2003categories}{abbott2005containers}{abbot2003categoriesthesis}, but since such an indexed
family of sets can be represented by as a function
\[
  \begin{tikzcd}
    E \arrow[d, "\pi"] \\
    B
  \end{tikzcd}
\]
we will refer to it as a \emph{bundle}.

@@

\begin{thm}[Theorem 2.17 of \cite{kock2012polynomial}]
The category of polynomial functors and natural transformations is equivalent to
the category of bundles and contravariant morphisms.
\end{thm}

@@

We will show in this paper that Dirichlet functors are, quite robustly, the
contravariant analogue of polynomial functors. In particular, the many equivalent ways to
say that a functor is polynomial have contravariant analogues.
\begin{thm}[See discussion in 1.18 of \cite{kock2012polynomial}]\label{thm:polynomial.set.characterization}
  Let $P : \Set \to \Set$ be a functor. Then the following are equivalent.
  \begin{enumerate}
  \item $P$ is polynomial.
  \item $P$ is the sum of covariant representables.
  \item There is a bundle $\pi : E \to B$ together with a natural isomorphism
    $$P(X) \cong \sum_{b \in B} X^{E_b}.$$
    Or, equivalently, a natural isomorphism of $P$ with the composite
    $$\Set \xto{{\Delta_{!_E}}} \Set_{/E} \xto{{\Pi_\pi}} \Set_{/B} \xto{{\Sigma_{!_B}}} \Set.$$
  \item $P$ is accessible and preserves connected limits.
  \end{enumerate}
\end{thm}

@@

Now, object-wise, a Dirichlet functor and a polynomial functor are determined by
the same data --- a bundle $\pi : E \to B$ of sets. Accordingly, one would
expect for any set $N$ a transformation
$$X^N \mapsto N^X$$
turning polynomial functors into Dirichlet functors, and vice versa. But the
natural transformations between each sort of functor induce different morphisms
between the bundles; natural transformations between polynomial functors induce
contravariant bundle morphisms, while natural transformations between Dirichlet
functors induce covariant bundle morphisms. However, if we restrict to those
morphisms of bundles which are \emph{isovariant} on the fibers --- that is, the
pullback diagrams of the form
\[
  \begin{tikzcd}[column sep=large]
E \arrow[r, "\term{tot}(f_{\sharp})"] \arrow[d, "\pi"'] \arrow[dr, phantom,
"\ulcorner" very near start] & E' \arrow[d, "\pi'"] \\
B \arrow[r, "f"']                                       & B'                  
\end{tikzcd}
\]
which preserve the number of elements in each row --- we find that such a morphism
is both a co- and contravariant morphism of bundles. It is well known that such
\emph{cartesian} morphisms of bundles correspond to \emph{cartesian} natural
transformations between polynomial functors \cite[Theorem 3.8]{kock2012polynomial} --- those whose naturality
squares are pullbacks. This is true as well for Dirichlet functors.
\begin{thm}
A natural transformation $D \to D'$ of Dirichlet functors is Cartesian if and
only if the corresponding bundle map
\[
  \begin{tikzcd}[column sep=large]
E \arrow[r, "\term{tot}(f_{\sharp})"] \arrow[d, "\pi"'] \arrow[dr, phantom,
"\ulcorner" very near start] & E' \arrow[d, "\pi'"] \\
B \arrow[r, "f"']                                       & B'                  
\end{tikzcd}
\]
is a pullback. As a corollary, we have an equivalence of categories
$$\type{Poly}_{\ulcorner} \simeq \Dir_{\ulcorner}$$
between polynomial functors with cartesian natural transfromations and Dirichlet
functors with cartesian natural transfromations.
\end{thm}

@@

\begin{defn}
  We let $\down$ be the walking arrow --- the category $\term{dom} \to
  \term{cod}$ consisting of a single morphism. We denote by $\sigma : \down
  \to (\downarrow)\op$ the self-dualizing isomorphism of $\down$, and note
  that $\sigma\inv = \sigma\op$.
\end{defn}

@@

The existence of the final adjoint can be deduced from the fact that $!^{(-)}$ preserves all
limits. It sends a bundle $\pi : E \to B$ to the largest subset $\term{ZC}(\pi)$ of $B$ for
which the following square is a pullback
\[
  \begin{tikzcd}[column sep=large]
0 \arrow[r] \arrow[d] \arrow[dr, phantom,
"\ulcorner" very near start] & E \arrow[d, "\pi"] \\
\term{ZC}(\pi) \arrow[r, hookrightarrow]                                       & B
\end{tikzcd}
\qedhere
\]
\end{proof}
We prove a few quick facts we will use later.
\begin{lem}\label{lem:bang.preserves.connected.colimits}
  The functor $!_{(-)}\colon\Set\to\Set^{\down}$ from \eqref{eqn.bangbang} preserves connected colimits.
\end{lem}
\begin{proof}
  To see that $!_{(-)}$ preserves connected colimits, recall that colimits in $\Set^{\down}$ are calculated pointwise. It remains to show, then, that the map of bundles
  \[
    \begin{tikzcd}
    \colim X_i \arrow[r, equals] \arrow[d] & \colim X_i \arrow[d] \\
    \colim 1 \arrow[r] & 1 
    \end{tikzcd}
  \]
  is an isomorphism in $\Set^{\down}$, for which it suffices to show that
  $\colim 1$ is terminal. But the colimit of a diagram of terminal objects is the set
  of connected components of its indexing category. Since we assumed the
  indexing category was connected, this contains a single element.
\end{proof}

@@

\begin{rmk}
 Since polynomial functors preserve connected limits, the composite $P \circ D$
 of a polynomial functor after a Dirichlet functor is Dirichlet. On the other
 hand, the composite $D' \circ D\op$ of two Dirichlet functors does not in
 general preserve connected limits, since $D\op$ sends connected colimits in
 $\Set$ to connected colimits in $\Set\op$, and $D'$ does not necessarily
 preserve these. Furthermore, composites of Dirichlet functors are not in
 general accessible.   
\end{rmk}

\begin{rmk}
The six adjoints of \cref{prop:adjoint.sextuple} correspond, under
the equivalence of \cref{thm:dirichlet.set.characterization}, to
\begin{align*}
  \term{ZC}(D!_0) &\cong \mbox{the coefficient of $0^X$ in $D$.}\\
  \ext(!^{C}) &\cong X \mapsto C \times 0^X \\
  \cod(D!_0) &\cong D(0) \\
  \ext( \const(C) ) &\cong X \mapsto C \\
  \dom(D!_0) &\cong D(1) \\
  \ext( !_{C} ) &\cong X \mapsto C^X 
\end{align*}
 In particular, $!_{(-)}$ corresponds to the Yoneda embedding.
\end{rmk}

@@

\begin{thm}\label{thm:set.poly.cart.equiv.dir.cart}
  There is an equivalence of categories
  $$\type{Poly}_{\ulcorner} \simeq \Dir_{\ulcorner}$$
  between the category of polynomial functors and cartesian transformations and
  Dirichlet functors and cartesian transformations. This equivlalence sends
  representables to representables
  $$(-)^N \mapsto N^{(-)}.$$
\end{thm}
\begin{proof}
This follows by composing the equivalence of $\Dir_{\ulcorner}$ with
$\Set^{\down}$ from
\cref{prop:set.dirichlet.cartesian.iff.bundle.cartesian} with that of
\cite[Proposition 3.14]{kock2012polynomial}, noting that
$\Set^{\down}_{\ulcorner}$ is the category of $(1,1)$-polynomials. 
\end{proof}

\begin{cor}
  Let $D$ be a Dirichlet functor. Then the category
  $\Dir_{\ulcorner/D}$
  of Dirichlet functors with a cartesian map to $D$ is a topos.
\end{cor}
\begin{proof}
By Theorem \ref{thm:set.poly.cart.equiv.dir.cart}, this category is equivalent
to $\type{Poly}_{\ulcorner/P}$ for a polynomial $P$. But this is a topos as
observed in \cite[Remark 2.6.2]{GHK:Analytic.Monads}.
\end{proof}

@@

  with left exact left adjoint given by the vertical / cartesian factorization system:
  \[
\begin{tikzcd}
E' \arrow[r] \arrow[d] & E \arrow[d, "\pi"] \\
B' \arrow[r]           & B                 
\end{tikzcd} \quad\quad\mapsto\quad\quad
\begin{tikzcd}
E' \arrow[d] \arrow[r, dashed]    & \bullet \arrow[d] \arrow[r] \arrow[rd, phantom, "\ulcorner" very near start] & E \arrow[d, "\pi"] \\
B' \arrow[r, equals] & B' \arrow[r]                           & B                 
\end{tikzcd}
  \]
  As a corollary, $\Dir_{\ulcorner/D} \hookrightarrow \Dir_{D}$ is a
  subtopos inclusion.
\end{thm}
\begin{proof}
We have displayed both the action of the left adjoint and its unit --- the
universal map into the pullback. The
counit is always an isomorphism since pullbacks are unique up to unique
isomorphism.
